---
title: "Discrete Choice Experiment — design, piloting and estimation"
subtitle: "Gender-Responsive Childcare Services in Uasin Gishu County Markets"
author: "Study team"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
  word_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
set.seed(20260918)   # fix the seed: the design must be reproducible
```

# What this file does

This is the full pipeline for the choice experiment in Section F of the trader
questionnaire, in the order it has to be run:

1. Define the seven attributes and their levels.
2. Generate a **pilot design** using zero priors.
3. Print the cards in the wording enumerators read aloud.
4. After the pilot (n = 30), estimate a model and take the coefficients as priors.
5. Regenerate a **Bayesian D-efficient design** using those priors — this is the
   version the 300 respondents see.
6. Report willingness to pay from the final model.

Steps 4 and 5 are the ones that get dropped under time pressure. If they are
dropped, the write-up has to say the design used assumed priors.

# Packages

```{r packages, eval=FALSE}
install.packages(c("idefix", "mlogit", "dfidx", "dplyr", "tidyr", "knitr"))
```

Requires **R 4.1 or later** — the code uses the native pipe `|>`. The chunks
above `# After the pilot` that are not marked `eval=FALSE` run on knit, so
`idefix`, `mlogit` and `knitr` must be installed before you knit this file.

```{r libs}
library(idefix)
library(dplyr)
library(tidyr)
library(knitr)
```

# Attributes and levels

Seven attributes. The design note in an earlier draft said six; the cards
carried seven. Seven is correct.

```{r attributes}
attr_levels <- list(
  cost      = c("KES 30/day", "KES 60/day", "KES 100/day", "KES 150/day"),
  location  = c("Inside the market, next to the stalls",
                "5-minute walk", "15-minute walk"),
  hours     = c("7:00-17:00", "6:00-19:00"),
  care      = c("1 trained caregiver per 5 children",
                "1 trained caregiver per 10 children",
                "1 untrained helper per 10 children"),
  food      = c("Porridge and lunch provided", "You bring the child's food"),
  operator  = c("County staff; complaints go to the county office",
                "Private operator paying rent; operator sets fees",
                "County and private operator jointly agree fees",
                "Committee of market traders; complaints go to the committee"),
  inclusion = c("No special accommodation for children with disabilities",
                "Physically accessible (ramp, adapted toilet), staff not trained",
                paste("Physically accessible with staff trained to support",
                      "children with disabilities"))
)

lvls   <- lengths(attr_levels)          # c(4, 3, 2, 3, 2, 4, 3)
coding <- rep("D", length(lvls))        # dummy coding for all seven
lvls
```

**Before this design is finalised**, confirm with the County (regulatory
interview, section 1) whether *"1 untrained helper per 10 children"* and
*"no special accommodation"* describe arrangements that could actually be
licensed. If they could not, drop those levels here and regenerate — otherwise
we are pricing a service nobody may legally offer.

```{r drop-unlawful, eval=FALSE}
# Run this ONLY if the County confirms those levels are below the enforced minimum.
attr_levels$care      <- attr_levels$care[1:2]
attr_levels$inclusion <- attr_levels$inclusion[2:3]
lvls   <- lengths(attr_levels)
coding <- rep("D", length(lvls))
```

## How many parameters, and how many cards

```{r parameters}
n_par  <- sum(lvls - 1) + 1     # dummy-coded main effects + one no-choice constant
n_sets <- 12                    # total choice sets, split into 2 blocks of 6
n_alts <- 3                     # A, B, and "neither"

data.frame(
  quantity = c("Parameters to estimate", "Choice sets", "Sets per respondent",
               "Minimum sets needed (params / (alts - 1))"),
  value    = c(n_par, n_sets, n_sets / 2, ceiling(n_par / (n_alts - 1)))
) |> kable()
```

Twelve sets clears the identification minimum. If the pilot shows the cards are
quick to administer, raising to 16 sets (two blocks of 8) buys precision
cheaply — worth revisiting after the pilot rather than deciding now.

# Pilot design — zero priors

With no data yet, the priors are zero: we are not claiming to know which way
traders lean, only building a design that is efficient if they lean nowhere.

```{r pilot-design, eval=FALSE}
# One prior per parameter. The no-choice constant is passed separately.
mu_nochoice <- 0
mu_attrs    <- rep(0, sum(lvls - 1))

# For a fixed (non-Bayesian) prior, par.draws is a single row.
# With alt.cte, idefix expects a LIST: constants first, then attribute parameters.
par_draws <- list(
  matrix(mu_nochoice, nrow = 1),
  matrix(mu_attrs,    nrow = 1)
)

pilot <- CEA(
  lvls      = lvls,
  coding    = coding,
  n.alts    = n_alts,
  n.sets    = n_sets,
  par.draws = par_draws,
  alt.cte   = c(0, 0, 1),   # constant on the opt-out only
  no.choice = 3,            # alternative 3 is "I'd keep my current arrangement"
  n.start   = 12,
  parallel  = TRUE
)

pilot$error          # D-error: lower is better, compare across restarts
saveRDS(pilot, "pilot_design.rds")
```

`CEA()` builds sets from the attribute levels directly. `Modfed()` is the
alternative — it searches a candidate set from `Profiles()` and is slower here
because the full factorial is `r prod(c(4,3,2,3,2,4,3))` profiles.

> **Check the argument names against your installed version** (`?CEA`) before
> the first run. `idefix`'s signature for `par.draws` and `alt.cte` has changed
> between releases, and this is the step most likely to need adjusting.

# Turning the design into readable cards

```{r decode, eval=FALSE}
pilot <- readRDS("pilot_design.rds")

decoded <- Decode(
  des       = pilot$design,
  n.alts    = n_alts,
  lvl.names = attr_levels,
  coding    = coding,
  alt.cte   = c(0, 0, 1),
  no.choice = 3
)

cards <- decoded$design |>
  as.data.frame() |>
  setNames(names(attr_levels)) |>
  mutate(
    set = rep(seq_len(n_sets), each = n_alts),
    alt = rep(c("Alternative A", "Alternative B", "Neither"), times = n_sets)
  ) |>
  filter(alt != "Neither")        # the opt-out is read aloud, not printed as a column

write.csv(cards, "choice_cards_long.csv", row.names = FALSE)
```

## Blocking into two versions

Each respondent answers six sets, not twelve. Odd respondent IDs get Block 1,
even IDs get Block 2. Blocks should be balanced — each level should appear
about as often in one block as the other. The draft design failed this: the
trader-committee option appeared on only two of twelve cards.

```{r blocking, eval=FALSE}
# Try many random splits, keep the one where level frequencies differ least.
balance_score <- function(idx, cards) {
  b1 <- cards |> filter(set %in% idx)
  b2 <- cards |> filter(!set %in% idx)
  sum(sapply(names(attr_levels), function(a) {
    t1 <- table(factor(b1[[a]], levels = attr_levels[[a]]))
    t2 <- table(factor(b2[[a]], levels = attr_levels[[a]]))
    sum(abs(t1 - t2))
  }))
}

candidates <- replicate(5000, sort(sample(seq_len(n_sets), n_sets / 2)),
                        simplify = FALSE)
scores     <- sapply(candidates, balance_score, cards = cards)
block1     <- candidates[[which.min(scores)]]

cards <- cards |> mutate(block = ifelse(set %in% block1, 1, 2))
cat("Best imbalance score:", min(scores), " (0 = perfectly balanced)\n")

# Check it by eye before accepting
cards |> count(block, operator) |> pivot_wider(names_from = block,
                                               values_from = n) |> kable()
```

## Printing the cards

```{r print-cards, eval=FALSE}
print_card <- function(d, s) {
  rows <- d |> filter(set == s)
  data.frame(
    Attribute       = c("Cost", "Location", "Opening hours",
                        "Caregiver ratio & training", "Food", "Who runs it",
                        "Inclusiveness of facilities"),
    `Alternative A` = unlist(rows[1, names(attr_levels)]),
    `Alternative B` = unlist(rows[2, names(attr_levels)]),
    check.names     = FALSE
  )
}

for (s in sort(unique(cards$set))) {
  cat("\n\n### Set", s, "(block", cards$block[cards$set == s][1], ")\n\n")
  print(kable(print_card(cards, s)))
  cat("\nWhich would you choose?  [ ] A   [ ] B   ",
      "[ ] Neither - I'd keep my current arrangement\n")
}
```

Knit to `word_document` and the cards come out formatted for printing. The
show-cards the respondents see also need **icons**, not only text — some
traders will not read comfortably. Budget the illustration time.

# After the pilot — estimate priors

Run this on the 30 pilot interviews. The point is not the results; it is to get
coefficients to feed back into the design.

```{r estimate, eval=FALSE}
library(mlogit); library(dfidx)

# Expected long format: one row per alternative per choice task.
# chid = unique task id, alt = 1/2/3, choice = TRUE for the chosen alternative.
pilot_data <- read.csv("pilot_responses_long.csv") |>
  mutate(across(c(cost, location, hours, care, food, operator, inclusion),
                as.factor),
         nochoice = as.integer(alt == 3))

idx <- dfidx(pilot_data, idx = c("chid", "alt"), choice = "choice")

m0 <- mlogit(choice ~ cost + location + hours + care + food +
               operator + inclusion + nochoice | 0, data = idx)
summary(m0)

priors <- coef(m0)
saveRDS(priors, "pilot_priors.rds")
```

## Diagnostics the pilot has to pass

```{r diagnostics, eval=FALSE}
resp <- read.csv("pilot_responses_long.csv")

# Always chose the cheapest option? (flagged on the cover sheet as F7)
resp |> filter(choice) |> group_by(respondent_id) |>
  summarise(always_cheapest = all(cost == min(cost)), .groups = "drop") |>
  count(always_cheapest)

# Always chose "neither"? (F8)
resp |> filter(choice) |> group_by(respondent_id) |>
  summarise(always_optout = all(alt == 3), .groups = "drop") |>
  count(always_optout)

# Left-right bias: A and B should be chosen about equally often
resp |> filter(choice, alt != 3) |> count(alt)
```

If a large share always pick the cheapest option, the price range is probably
too high — check it against what traders report earning (B22 and B23) and what
existing daycares charge, then adjust the levels before the main design.

# Final design — Bayesian, with pilot priors

```{r final-design, eval=FALSE}
priors <- readRDS("pilot_priors.rds")

# Draw from the estimated distribution rather than using point estimates:
# the design is then efficient across the uncertainty, not just at the mean.
n_draws <- 100
draws <- MASS::mvrnorm(n_draws, mu = priors, Sigma = vcov(m0))

par_draws_final <- list(
  matrix(draws[, "nochoice"], ncol = 1),
  draws[, setdiff(colnames(draws), "nochoice"), drop = FALSE]
)

final <- CEA(
  lvls = lvls, coding = coding,
  n.alts = n_alts, n.sets = n_sets,
  par.draws = par_draws_final,
  alt.cte = c(0, 0, 1), no.choice = 3,
  n.start = 20, parallel = TRUE
)

final$error
saveRDS(final, "final_design.rds")
```

Then re-run the decode, blocking and printing chunks above against
`final_design.rds`. **This** is the version that goes to 300 respondents.

# Willingness to pay

```{r wtp, eval=FALSE}
m <- readRDS("main_model.rds")
cf <- coef(m)

# Cost as a factor gives a coefficient per level. For a money value per
# attribute, refit with cost entered as a number (its KES value).
cost_numeric <- c(30, 60, 100, 150)
main_data <- main_data |> mutate(cost_kes = cost_numeric[as.integer(cost)])

m_wtp <- mlogit(choice ~ cost_kes + location + hours + care + food +
                  operator + inclusion + nochoice | 0, data = idx_main)

b_cost <- coef(m_wtp)["cost_kes"]
wtp <- -coef(m_wtp) / b_cost        # KES per day for each level vs its reference
round(wtp[names(wtp) != "cost_kes"], 1)
```

Each figure is what the average trader would pay per day, in shillings, for
that feature relative to the reference level. Two cautions for the write-up:

- **Report a confidence interval, not just the point estimate.** WTP is a ratio
  of two estimates, so its uncertainty is wider than either. Use the delta
  method or bootstrap the ratio.
- **This is individual willingness to pay, not household.** Question G3 records
  whether a spouse must agree before the household pays. Cross-tabulate before
  treating any of these numbers as a price the household would actually hand
  over.

# Reproducing this

```{r session, eval=FALSE}
sessionInfo()
```

Commit `pilot_design.rds`, `pilot_priors.rds` and `final_design.rds` alongside
this file. The seed at the top fixes the design, but only for the same package
versions — record them.
