---
title: "Discrete Choice Experiment — design, piloting and estimation"
subtitle: "Gender-Responsive Childcare Services in Uasin Gishu County Markets"
author: "Study team"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
  word_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

# Read this first

**Knitting this file generates the design.** It produces the choice cards,
checks them, and writes `choice_cards.csv`. Sections 1 to 5 run; sections 6 to 8
need the pilot responses and stay switched off until those exist.

**No packages are required to generate the design.** Everything in
`dce_design_core.R` is base R. There is nothing to install and nothing that
breaks when a package changes its arguments between versions — which is what
made the earlier `idefix` version of this file fail twice. `idefix` remains a
perfectly good tool; it is simply not a dependency here.

Put `dce_design_core.R` in the same folder as this file, open this file in
RStudio, and press **Knit**.

The order the work has to happen in:

1. Define the seven attributes and their levels.
2. Generate the design and check it. *(runs on knit)*
3. Print the cards in the wording enumerators read aloud. *(runs on knit)*
4. After the pilot (n = 30), estimate a model and take its coefficients as priors.
5. Regenerate the design from those priors — the version 300 respondents see.
6. Report willingness to pay.

Steps 4 and 5 are the ones that get dropped under time pressure. If they are
dropped, the write-up has to say the design used assumed priors.

# The method, in one paragraph

Priors are zero — we are not claiming to know which way traders lean. That
makes every alternative equally likely, so the multinomial logit information
matrix is exact rather than approximated:
`I = Σ_s X_s' (diag(p) − p p') X_s` with `p = 1/3`. Maximising `det(I)` is
therefore a well-defined optimisation, and the D-error it reports is a real
number you can compare between designs, not a black box. The search is
coordinate exchange: start from a random design, and repeatedly try every level
of every attribute of every alternative, keeping any change that improves the
objective, until nothing improves. Several random starts, best one kept.

```{r source}
source("dce_design_core.R")
```

# Attributes and levels

Seven attributes. An earlier draft of the design note said six; the cards
carried seven. Seven is correct.

```{r attributes}
attrs <- dce_attributes()
data.frame(
  attribute = names(attrs),
  levels    = vapply(attrs, function(a) length(a$levels), 1L),
  ordered   = vapply(attrs, function(a) a$ordered, TRUE),
  row.names = NULL
) |> knitr::kable()
```

`ordered = TRUE` means more is plainly better (or plainly worse) — cheaper is
better, closer is better, a trained caregiver beats an untrained one. That
ordering is used to rule out cards where one option beats the other on
everything, which ask the respondent for no trade-off at all. "Who runs it"
has no such ordering, which is the point of asking about it.

**Before the design is finalised**, confirm with the County (regulatory
interview, section 1) whether *"1 untrained helper per 10 children"* and
*"no special accommodation"* describe arrangements that could be licensed. If
not, drop them here and re-knit — otherwise we are pricing a service nobody may
lawfully offer.

```{r drop-unlawful}
# Uncomment ONLY once the County confirms these fall below the enforced minimum.
#
# attrs$care$levels      <- attrs$care$levels[1:2]
# attrs$inclusion$levels <- attrs$inclusion$levels[1:2]
```

# Generating the design

How many cards are needed is not a free choice. The model has one parameter per
non-reference level, plus a constant for the opt-out; a design cannot have fewer
choice sets than parameters. Respondent burden is held at six cards each, and
the number of *blocks* flexes to satisfy that.

```{r size}
lvls  <- vapply(attrs, function(a) length(a$levels), 1L)
n_par <- sum(lvls - 1) + 1
sets_per_respondent <- 6
n_blocks <- ceiling(n_par / sets_per_respondent)
n_sets   <- n_blocks * sets_per_respondent

data.frame(
  quantity = c("Parameters to estimate", "Minimum choice sets (= parameters)",
               "Cards seen per respondent", "Blocks", "Total cards generated"),
  value    = c(n_par, n_par, sets_per_respondent, n_blocks, n_sets)
) |> knitr::kable()
```

Three constraints are enforced during the search, not patched up afterwards:

- the two options on a card are never identical;
- neither option is weakly better than the other on every ordered attribute
  with the management model held equal;
- all seven attributes differ on every card, so each card asks for a real
  trade-off.

A small level-balance penalty is added to the objective so that ties in
D-efficiency break toward a design where each level appears equally often.
Without it a level can appear ten times and another eight, which makes even
blocks arithmetically impossible later.

```{r generate, cache=TRUE}
design <- dce_generate(
  attrs               = attrs,
  n_sets              = n_sets,
  sets_per_respondent = sets_per_respondent,
  n_start             = 10,      # raise for a slightly better design; costs time
  balance_weight      = 0.002,   # from a sweep: perfect balance for ~3% D-error
  seed                = 20260918,
  verbose             = FALSE
)
saveRDS(design, "dce_design.rds")
```

# Checking it before anyone prints it

```{r check}
dce_check(design)
```

Read this table before the cards go to a printer. `Identical A/B pairs` and
`Dominated pairs` must both be zero. `Attributes differing` should be 7 on every
card. Each block should carry the four management models an equal number of
times — that is the attribute the whole study exists to compare, and a model
that appears twice in a block cannot be estimated from it.

```{r write-csv}
cards <- dce_cards(design)
write.csv(cards, "choice_cards.csv", row.names = FALSE)
cat("wrote choice_cards.csv —", nrow(cards), "rows\n")
```

## Which block does a respondent see?

Each respondent answers ONE block. Assign by respondent ID:
`block = ((ID - 1) %% n_blocks) + 1`. Record the block on the cover sheet
(question F10) — the analysis cannot proceed without knowing which cards the
respondent saw.

```{r block-table}
data.frame(
  `Respondent ID` = paste0("ID mod ", n_blocks, " = ",
                           c(seq_len(n_blocks - 1), 0)),
  Block           = seq_len(n_blocks),
  check.names     = FALSE
) |> knitr::kable()
```

# The cards

```{r cards, results='asis'}
invisible(dce_print_cards(design, markdown = TRUE))
```

Knit to `word_document` and these come out formatted for printing. The
show-cards respondents actually see also need **icons**, not only text — some
traders will not read comfortably. Budget the illustration time.

# After the pilot — estimate priors

Everything below needs the pilot responses. Switch `eval=FALSE` to `eval=TRUE`
once `pilot_responses_long.csv` exists. These sections do need packages:
`install.packages(c("mlogit", "dfidx"))`.

```{r estimate, eval=FALSE}
library(mlogit); library(dfidx)

# One row per alternative per choice task. chid = task id, alt = 1/2/3,
# choice = TRUE for the alternative the respondent picked.
pilot_data <- read.csv("pilot_responses_long.csv") |>
  transform(nochoice = as.integer(alt == 3))
for (v in names(dce_attributes()))
  pilot_data[[v]] <- factor(pilot_data[[v]])

idx <- dfidx(pilot_data, idx = c("chid", "alt"), choice = "choice")
m0  <- mlogit(choice ~ cost + location + hours + care + food +
                operator + inclusion + nochoice | 0, data = idx)
summary(m0)
saveRDS(list(coef = coef(m0), vcov = vcov(m0)), "pilot_priors.rds")
```

## Diagnostics the pilot has to pass

```{r diagnostics, eval=FALSE}
resp <- read.csv("pilot_responses_long.csv")
chosen <- resp[resp$choice, ]

# Always picked the cheapest option? (F7 on the cover sheet)
table(tapply(chosen$cost, chosen$respondent_id, function(x) all(x == min(x))))

# Always picked "neither"? (F8)
table(tapply(chosen$alt, chosen$respondent_id, function(x) all(x == 3)))

# Left-right bias: A and B should be picked about equally often
table(chosen$alt[chosen$alt != 3])
```

If a large share always pick the cheapest option, the price range is probably
too high — check it against what traders report earning (B22 and B23) and what
existing daycares charge, then adjust the levels and regenerate.

# Final design — using the pilot priors

The generator above uses zero priors. Once real priors exist, the efficient
design is the one that minimises the D-error *at those priors*. That is a
one-line change to `.d_error()` — replace the uniform `p` with the choice
probabilities implied by the estimated coefficients — or the point at which
`idefix` earns its keep, since its Bayesian routines do exactly this.

```{r final-design, eval=FALSE}
pri <- readRDS("pilot_priors.rds")

# Option A — idefix, if it is installed and you want Bayesian priors:
# library(idefix)
# draws <- MASS::mvrnorm(100, mu = pri$coef, Sigma = pri$vcov)
# cte   <- grep("nochoice", colnames(draws))
# final <- CEA(lvls = lvls, coding = rep("D", length(lvls)),
#              n.sets = n_sets, n.alts = 3,
#              par.draws = list(draws[, cte, drop = FALSE],
#                               draws[, -cte, drop = FALSE]),
#              alt.cte = c(0, 0, 1), no.choice = TRUE,   # logical here
#              n.blocks = n_blocks, n.start = 20)
#
# Option B — keep this file's generator and weight by the estimated priors.
# See the note above .d_error() in dce_design_core.R.
```

# Willingness to pay

```{r wtp, eval=FALSE}
cost_numeric <- c(30, 60, 100, 150)
main_data <- read.csv("main_responses_long.csv") |>
  transform(cost_kes = cost_numeric[as.integer(factor(cost))],
            nochoice = as.integer(alt == 3))
idx_main <- dfidx(main_data, idx = c("chid", "alt"), choice = "choice")

m_wtp <- mlogit(choice ~ cost_kes + location + hours + care + food +
                  operator + inclusion + nochoice | 0, data = idx_main)

b_cost <- coef(m_wtp)["cost_kes"]
round(-coef(m_wtp)[names(coef(m_wtp)) != "cost_kes"] / b_cost, 1)
```

Each figure is what the average trader would pay per day, in shillings, for that
feature relative to its reference level. Two cautions for the write-up:

- **Report an interval, not just the point estimate.** WTP is a ratio of two
  estimates, so its uncertainty is wider than either. Use the delta method or
  bootstrap the ratio.
- **This is individual willingness to pay, not household.** Question G3 records
  whether a spouse must agree before the household pays. Cross-tabulate before
  treating any of these as a price a household would actually hand over.

# Reproducing this

```{r session}
sessionInfo()
```

The seed in the `generate` chunk fixes the design. Commit `dce_design.rds` and
`choice_cards.csv` alongside this file so the printed cards can always be traced
back to the design that produced them.
