---
title: "Discrete Choice Experiment — design, piloting and estimation"
subtitle: "Gender-Responsive Childcare Services in Uasin Gishu County Markets"
author: "Study team"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
  word_document:
    toc: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
set.seed(20260918)   # fix the seed: the design must be reproducible
```

# Read this first

**Sections 1 to 5 run when you knit this file.** Knitting produces a real
D-efficient design, writes `choice_cards_long.csv`, and prints all twelve cards
ready to read to a respondent. You do not have to run anything by hand.

**Sections 6 to 8 are switched off** (`eval=FALSE`) because they need the pilot
responses, which do not exist yet. Once `pilot_responses_long.csv` is on disk,
change `eval=FALSE` to `eval=TRUE` in those chunks and knit again.

The order the work has to happen in:

1. Define the seven attributes and their levels.
2. Generate a **pilot design** using zero priors. *(runs on knit)*
3. Print the cards in the wording enumerators read aloud. *(runs on knit)*
4. After the pilot (n = 30), estimate a model and take its coefficients as priors.
5. Regenerate a **Bayesian D-efficient design** from those priors — this is the
   version the 300 respondents see.
6. Report willingness to pay from the final model.

Steps 4 and 5 are the ones that get dropped under time pressure. If they are
dropped, the write-up has to say the design used assumed priors.

# Packages

```{r packages, eval=FALSE}
install.packages(c("idefix", "mlogit", "dfidx", "dplyr", "tidyr", "knitr", "MASS"))
```

Requires **R 4.1 or later** — the code uses the native pipe `|>`. `idefix`,
`dplyr`, `tidyr` and `knitr` must be installed before you knit.

```{r libs}
library(idefix)
library(dplyr)
library(tidyr)
library(knitr)
```

# Attributes and levels

Seven attributes. An earlier draft of the design note said six; the cards
carried seven. Seven is correct.

```{r attributes}
attr_levels <- list(
  cost      = c("KES 30/day", "KES 60/day", "KES 100/day", "KES 150/day"),
  location  = c("Inside the market, next to the stalls",
                "5-minute walk", "15-minute walk"),
  hours     = c("7:00-17:00", "6:00-19:00"),
  care      = c("1 trained caregiver per 5 children",
                "1 trained caregiver per 10 children",
                "1 untrained helper per 10 children"),
  food      = c("Porridge and lunch provided", "You bring the child's food"),
  operator  = c("County staff; complaints go to the county office",
                "Private operator paying rent; operator sets fees",
                "County and private operator jointly agree fees",
                "Committee of market traders; complaints go to the committee"),
  inclusion = c("No special accommodation for children with disabilities",
                "Physically accessible (ramp, adapted toilet), staff not trained",
                paste("Physically accessible with staff trained to support",
                      "children with disabilities"))
)

lvls   <- lengths(attr_levels)          # c(4, 3, 2, 3, 2, 4, 3)
coding <- rep("D", length(lvls))        # dummy coding for all seven
lvls
```

**Before this design is finalised**, confirm with the County (regulatory
interview, section 1) whether *"1 untrained helper per 10 children"* and
*"no special accommodation"* describe arrangements that could actually be
licensed. If they could not, uncomment the chunk below and re-knit — otherwise
we are pricing a service nobody may legally offer.

```{r drop-unlawful}
# Switch eval to TRUE (or uncomment) ONLY once the County confirms these levels
# fall below the enforced minimum standard.
#
# attr_levels$care      <- attr_levels$care[1:2]
# attr_levels$inclusion <- attr_levels$inclusion[2:3]
# lvls   <- lengths(attr_levels)
# coding <- rep("D", length(lvls))
```

## How many parameters, and how many cards

`idefix` refuses to build a design with fewer choice sets than parameters. That
is stricter than the textbook rule (`parameters / (alternatives - 1)`), and it
is the binding constraint here — so derive the number of sets from it rather
than picking a round number.

Respondent burden is held fixed at six sets each; the number of *blocks* flexes
to satisfy identification. Keeping six sets per person matters more than keeping
two blocks: these are stall-side interviews in a working market.

```{r parameters}
n_alts <- 3    # Alternative A, Alternative B, and the opt-out.
               # n.alts COUNTS the no-choice alternative.
n_cte  <- 1    # one constant, on the opt-out
n_par  <- sum(lvls - 1) + n_cte

sets_per_respondent <- 6
n_blocks <- ceiling(n_par / sets_per_respondent)   # enough sets to identify
n_sets   <- n_blocks * sets_per_respondent

# idefix stops with "Model is unidentified" if this is not met.
stopifnot(n_sets >= n_par)

data.frame(
  quantity = c("Parameters to estimate",
               "Choice sets idefix requires (>= parameters)",
               "Sets per respondent",
               "Blocks needed",
               "Total choice sets generated"),
  value    = c(n_par, n_par, sets_per_respondent, n_blocks, n_sets)
) |> kable()
```

With the full seven attributes this comes out at **three blocks of six**, so
each respondent still answers six cards and no one sees more than that.

If you would rather keep two blocks, set `sets_per_respondent <- 8` instead:
that gives two blocks of eight, which also satisfies identification but asks
each trader for two more cards. Dropping the unlawful levels (the chunk above)
lowers the parameter count and may bring the design back within two blocks —
re-knit and read the table.

**This changes the block-assignment instruction in Section F of the
questionnaire**, which currently says odd respondent IDs get Block 1 and even
IDs Block 2. With three blocks it becomes `block = ((ID - 1) %% 3) + 1`. The
printed cards below carry their block number, so use those.

# Pilot design — zero priors

With no data yet the priors are zero: we are not claiming to know which way
traders lean, only building a design that is efficient if they lean nowhere.

Three details in this call are easy to get wrong, so they are worth naming:

- `n.alts` **includes** the opt-out, so it is 3, not 2.
- `alt.cte` must carry a `1` in the **last** position when `no.choice = TRUE`;
  idefix stops with an error otherwise.
- `no.choice` here is **`TRUE`**, a logical. In `Decode()` further down, the
  argument of the same name is an **integer** naming which alternative is the
  opt-out. They are not interchangeable.

```{r pilot-design, cache=TRUE}
# par.draws must be a LIST of two matrices when any alternative has a constant:
# first the constants, then the attribute parameters. Equal numbers of rows.
par_draws <- list(
  matrix(0, nrow = 1, ncol = n_cte),            # prior on the opt-out constant
  matrix(0, nrow = 1, ncol = sum(lvls - 1))     # priors on the attributes
)

pilot <- CEA(
  lvls          = lvls,
  coding        = coding,
  n.sets        = n_sets,
  n.alts        = n_alts,
  par.draws     = par_draws,
  alt.cte       = c(0, 0, 1),
  no.choice     = TRUE,
  n.blocks      = n_blocks,   # idefix blocks the design itself
  blocking.iter = 50,
  n.start       = 5,      # raise for a better design once you are past testing
  parallel      = FALSE   # keeps the knit reproducible under the seed above
)

saveRDS(pilot, "pilot_design.rds")
cat("DB-error (lower is better):", pilot$DB.error, "\n")
```

## Is the design balanced?

`CEA()` reports this itself, which is more reliable than counting by hand. Every
level should appear a similar number of times; a level that barely appears
cannot be estimated precisely. The earlier draft design failed here — the
trader-committee option appeared on only two of twelve cards.

```{r balance}
pilot$level.count
cat("\nLevel overlap within choice sets:\n"); print(pilot$level.overlap)
```

# Turning the design into readable cards

## What CEA actually returned

`idefix` is not consistent between versions about whether the design matrix
includes rows for the opt-out alternative, or whether the alternative-specific
constant appears as a leading column. Rather than assume, look.

```{r design-shape}
des_raw <- as.matrix(pilot$design)
cat("design rows:", nrow(des_raw), " columns:", ncol(des_raw), "\n")
cat("choice sets:", n_sets, "\n")

rows_per_set <- nrow(des_raw) / n_sets
cat("rows per choice set:", rows_per_set,
    if (rows_per_set == n_alts) " (the opt-out has its own row)"
    else " (the opt-out is implied, not stored)", "\n")

n_attr_cols <- sum(lvls - 1)
cat("attribute columns expected:", n_attr_cols,
    "| leading constant columns:", ncol(des_raw) - n_attr_cols, "\n")

stopifnot(rows_per_set %% 1 == 0)
head(des_raw, 4)
```

## Decoding

`Decode()` needs `nrow(des)` to divide exactly by the `n.alts` you hand it, and
what it expects differs by version. Dummy coding is simple enough to invert
directly, so we do that instead and remove the version dependency: for an
attribute with L levels, idefix writes L-1 columns, level 1 is the reference
(all zeros), and level k has a 1 in column k-1.

```{r decode}
decode_design <- function(des, lvls, lvl_names) {
  des <- as.matrix(des)
  n_attr_cols <- sum(lvls - 1)
  # alternative-specific constants are written as leading columns — drop them
  if (ncol(des) > n_attr_cols) {
    des <- des[, (ncol(des) - n_attr_cols + 1):ncol(des), drop = FALSE]
  }
  stopifnot(ncol(des) == n_attr_cols)

  out <- matrix(NA_character_, nrow = nrow(des), ncol = length(lvls))
  j0 <- 1
  for (j in seq_along(lvls)) {
    k <- lvls[j] - 1
    blk <- des[, j0:(j0 + k - 1), drop = FALSE]
    idx <- apply(blk, 1, function(r) if (all(r == 0)) 1L else which.max(r) + 1L)
    out[, j] <- lvl_names[[j]][idx]
    j0 <- j0 + k
  }
  colnames(out) <- names(lvl_names)
  as.data.frame(out, stringsAsFactors = FALSE)
}

des <- decode_design(des_raw, lvls, attr_levels)

cards <- des |>
  mutate(
    set = rep(seq_len(n_sets), each = rows_per_set),
    alt = rep(seq_len(rows_per_set), times = n_sets)
  ) |>
  filter(alt <= 2) |>                                  # keep A and B; opt-out is read aloud
  mutate(alt = ifelse(alt == 1, "Alternative A", "Alternative B"))

# Blocks come from CEA, but versions differ on the shape. Accept it only if it
# really is one block label per choice set; otherwise split evenly.
blocks <- rep(seq_len(n_blocks), each = sets_per_respondent)   # fallback
if (!is.null(pilot$Blocks)) {
  b <- as.vector(unlist(pilot$Blocks))
  if (length(b) == n_sets && all(b %in% seq_len(n_blocks))) {
    blocks <- b
    cat("Using the block assignment returned by CEA.\n")
  } else {
    cat("CEA's Blocks component was not one label per set; split evenly instead.\n")
  }
}
cards$block <- blocks[cards$set]

write.csv(cards, "choice_cards_long.csv", row.names = FALSE)

cards |> count(block, operator) |>
  pivot_wider(names_from = block, values_from = n, names_prefix = "Block ",
              values_fill = 0) |>
  kable(caption = "Who-runs-it levels across the blocks — should be roughly even")
```

A sanity check worth reading before you print anything: every level should
appear, and no card should have two identical alternatives.

```{r card-checks}
stopifnot(nrow(cards) == n_sets * 2)

key <- apply(cards[names(attr_levels)], 1, paste, collapse = " | ")
same <- tapply(key, cards$set, function(k) length(unique(k)) == 1)
cat("Sets where A and B are identical:", sum(same), "(should be 0)\n")

missing_levels <- setdiff(unlist(attr_levels), unlist(cards[names(attr_levels)]))
if (length(missing_levels)) {
  cat("\nLevels that never appear on any card:\n")
  print(missing_levels)
} else cat("Every level appears at least once.\n")
```

## The cards

Each respondent answers one block. Assign by respondent ID:
`block = ((ID - 1) %% n_blocks) + 1`.

```{r print-cards, results='asis'}
attr_labels <- c("Cost", "Location", "Opening hours",
                 "Caregiver ratio & training", "Food", "Who runs it",
                 "Inclusiveness of facilities")

for (b in sort(unique(cards$block))) {
  cat("\n\n### Block", b, "\n")
  for (s in sort(unique(cards$set[cards$block == b]))) {
    rows <- cards |> filter(set == s)
    card <- data.frame(
      Attribute       = attr_labels,
      `Alternative A` = unlist(rows[rows$alt == "Alternative A", names(attr_levels)]),
      `Alternative B` = unlist(rows[rows$alt == "Alternative B", names(attr_levels)]),
      check.names     = FALSE, row.names = NULL
    )
    cat("\n\n**Set", which(sort(unique(cards$set[cards$block == b])) == s),
        "of", sets_per_respondent, "**  (design set", s, ")\n\n")
    print(kable(card))
    cat("\n\nWhich would you choose?",
        "&#9744; Alternative A &nbsp; &#9744; Alternative B &nbsp;",
        "&#9744; Neither — I'd keep my current arrangement\n")
  }
}
```

Knit to `word_document` and these come out formatted for printing. The
show-cards respondents actually see also need **icons**, not only text — some
traders will not read comfortably. Budget the illustration time.

# After the pilot — estimate priors

Everything from here needs the pilot responses. Switch `eval=FALSE` to
`eval=TRUE` once `pilot_responses_long.csv` exists.

```{r estimate, eval=FALSE}
library(mlogit); library(dfidx)

# Expected long format: one row per alternative per choice task.
# chid = unique task id, alt = 1/2/3, choice = TRUE for the chosen alternative.
pilot_data <- read.csv("pilot_responses_long.csv") |>
  mutate(across(c(cost, location, hours, care, food, operator, inclusion),
                as.factor),
         nochoice = as.integer(alt == 3))

idx <- dfidx(pilot_data, idx = c("chid", "alt"), choice = "choice")

m0 <- mlogit(choice ~ cost + location + hours + care + food +
               operator + inclusion + nochoice | 0, data = idx)
summary(m0)

saveRDS(list(coef = coef(m0), vcov = vcov(m0)), "pilot_priors.rds")
```

## Diagnostics the pilot has to pass

```{r diagnostics, eval=FALSE}
resp <- read.csv("pilot_responses_long.csv")

# Always chose the cheapest option? (flagged on the cover sheet as F7)
resp |> filter(choice) |> group_by(respondent_id) |>
  summarise(always_cheapest = all(cost == min(cost)), .groups = "drop") |>
  count(always_cheapest)

# Always chose "neither"? (F8)
resp |> filter(choice) |> group_by(respondent_id) |>
  summarise(always_optout = all(alt == 3), .groups = "drop") |>
  count(always_optout)

# Left-right bias: A and B should be chosen about equally often
resp |> filter(choice, alt != 3) |> count(alt)
```

If a large share always pick the cheapest option, the price range is probably
too high — check it against what traders report earning (B22 and B23) and what
existing daycares charge, then adjust the levels before the main design.

# Final design — Bayesian, with pilot priors

```{r final-design, eval=FALSE}
pri <- readRDS("pilot_priors.rds")

# Draw from the estimated distribution rather than using point estimates: the
# design is then efficient across the uncertainty, not only at the mean.
draws <- MASS::mvrnorm(100, mu = pri$coef, Sigma = pri$vcov)

cte_cols <- grep("nochoice", colnames(draws))
par_draws_final <- list(
  draws[, cte_cols, drop = FALSE],
  draws[, -cte_cols, drop = FALSE]
)

final <- CEA(
  lvls = lvls, coding = coding,
  n.sets = n_sets, n.alts = n_alts,
  par.draws = par_draws_final,
  alt.cte = c(0, 0, 1), no.choice = TRUE,
  n.blocks = n_blocks, blocking.iter = 50,
  n.start = 20, parallel = TRUE
)

cat("DB-error:", final$DB.error, "\n")
saveRDS(final, "final_design.rds")
```

Then re-run the decode and card chunks against `final_design.rds` by changing
`pilot` to `final` at the top of the decode chunk. **That** is the version the
300 respondents see.

# Willingness to pay

```{r wtp, eval=FALSE}
# Cost as a factor gives a coefficient per level. For a money value per
# attribute, refit with cost entered as a number (its KES value).
cost_numeric <- c(30, 60, 100, 150)
main_data <- read.csv("main_responses_long.csv") |>
  mutate(cost_kes = cost_numeric[as.integer(factor(cost))],
         nochoice = as.integer(alt == 3))
idx_main <- dfidx(main_data, idx = c("chid", "alt"), choice = "choice")

m_wtp <- mlogit(choice ~ cost_kes + location + hours + care + food +
                  operator + inclusion + nochoice | 0, data = idx_main)

b_cost <- coef(m_wtp)["cost_kes"]
wtp <- -coef(m_wtp) / b_cost        # KES per day for each level vs its reference
round(wtp[names(wtp) != "cost_kes"], 1)
```

Each figure is what the average trader would pay per day, in shillings, for
that feature relative to the reference level. Two cautions for the write-up:

- **Report a confidence interval, not just the point estimate.** WTP is a ratio
  of two estimates, so its uncertainty is wider than either. Use the delta
  method or bootstrap the ratio.
- **This is individual willingness to pay, not household.** Question G3 records
  whether a spouse must agree before the household pays. Cross-tabulate before
  treating any of these numbers as a price a household would actually hand over.

# Reproducing this

```{r session}
sessionInfo()
```

Commit `pilot_design.rds`, `pilot_priors.rds` and `final_design.rds` alongside
this file. The seed at the top fixes the design, but only for the same package
versions — which is what `sessionInfo()` above records.
