---
title: "Pilot follow-up — what the DCE is telling us, and what to change"
subtitle: "Gender-Responsive Childcare Services in Uasin Gishu County Markets"
author: "Study team"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output:
  word_document: { toc: true, number_sections: true }
  html_document: { toc: true, toc_float: true, number_sections: true }
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, comment = "")
```

# How to run this

Put `dce_long.csv`, `respondents.csv` and `pilot_dce_analysis.R` in the same
folder as this file and knit. Base R only — no packages to install. The
analysis script is the single source of truth for every number quoted below;
the narrative here reads its output rather than recomputing anything.

```{r run, results='hide'}
out <- capture.output(source("pilot_dce_analysis.R", echo = FALSE))
sec <- function(from, to = NULL) {
  i <- grep(from, out, fixed = TRUE)[1]
  j <- if (is.null(to)) length(out) else grep(to, out, fixed = TRUE)[1] - 1
  cat(paste(out[i:j], collapse = "\n"))
}
```

# What was fielded

```{r s1, results='asis'}
cat("```\n"); sec("== 1.", "== 2."); cat("\n```\n")
```

Twenty-three traders each completed all six choice tasks — a 100% task
completion rate, which is itself a good sign. Twelve saw Block 1, eleven saw
Block 2, so all twelve cards of the fielded design were used.

# The headline finding, and why the obvious reading of it is wrong

```{r s2, results='asis'}
cat("```\n"); sec("== 2.", "== 3."); cat("\n```\n")
```

Alternative A was chosen 59.1% of the time, which looks like a left-hand-side
bias. But the fielded cards put the **cheaper** option in position A on nine of
the twelve cards. Position and price therefore agree three quarters of the
time, and no amount of modelling can fully separate them in this data.

Note the second number: the cheaper option was chosen only 47.7% of the time —
a coin flip. If traders were answering on price, that figure would be high. It
is not. Price is not visibly driving these choices.

# Can the model be estimated at all?

```{r s3, results='asis'}
cat("```\n"); sec("== 3.", "== 4."); cat("\n```\n")
```

# The cost coefficient flips sign

```{r s4, results='asis'}
cat("```\n"); sec("== 4.", "== 5."); cat("\n```\n")
```

This is the single most important result in the pilot. Adding one position term
moves the price coefficient from negative to positive. A positive price
coefficient would mean traders prefer to pay more, which nobody believes. The
honest reading is that the data cannot tell price and position apart, so
**neither estimate should be used**.

# The full attribute model

```{r s5, results='asis'}
cat("```\n"); sec("== 5.", "== 6."); cat("\n```\n")
```

Not one service feature reaches significance. With twelve cards and fifteen
parameters that is expected, and it is not evidence that traders are
indifferent to these features — only that this pilot cannot measure them.

```{r s6, results='asis'}
cat("```\n"); sec("== 6.", "== 7."); cat("\n```\n")
```

The pseudo-R-squared looks respectable, but it is carried almost entirely by
the opt-out constant: traders almost always chose one of the two services
rather than nothing. It says little about the attributes.

# How individuals answered

```{r s7, results='asis'}
cat("```\n"); sec("== 7.", "== 8."); cat("\n```\n")
```

Four of twenty-three traders chose A on every single task. That is
straight-lining, not a preference.

# Interview duration

```{r s8, results='asis'}
cat("```\n"); sec("== 8.", "== 9."); cat("\n```\n")
```

# Reliability of the two scales

```{r s9, results='asis'}
cat("```\n"); sec("== 9.", "== 10."); cat("\n```\n")
```

```{r s10, results='asis'}
cat("```\n"); sec("== 10.", "== 11."); cat("\n```\n")
```

E1h has zero variance — every trader gave the same answer, so it cannot
correlate with anything and cannot belong to a scale. E6 is not a scale in
either direction: reverse-coding the two traditionally-worded items makes the
alpha collapse, and several item-total correlations turn negative. Both should
be analysed item by item.

# Sample size for the full study

```{r s11, results='asis'}
cat("```\n"); sec("== 11.", "== 12."); cat("\n```\n")
```

# Is the position finding robust?

```{r s12, results='asis'}
cat("```\n"); sec("== 12.", "== 13."); cat("\n```\n")
```

Remove the four straight-liners and the position effect is no longer
significant. So the problem is better described as a small group of traders not
engaging with the task, amplified by a design that put the cheaper option on
the same side most of the time. Both have fixes, and they are different fixes.

# Willingness to pay

```{r s13, results='asis'}
cat("```\n"); sec("== 13.", "== 14."); cat("\n```\n")
```

# Session information

```{r session, echo=TRUE}
sessionInfo()
```
